---
date: "2016-04-27T00:00:00Z"
external_link: ""
image:
  caption: Photo by rawpixel on Unsplash
  focal_point: Smart
summary: Museum of Modern Art Tour.
tags:
- Deep Learning
title: MoMA Lab
url_code: ""
url_pdf: ""
url_slides: ""
url_video: ""
---
# Workflow for Lab 02

The lab instructions can be found [here](https://stevenbedrick.github.io/data-vis-labs-2022/02-moma.html); we will work through its contents together via Webex. You will use this RMarkdown file as your workspace and final document. Don't forget to update the "author" metadata field at the top of the file!

# Start by Loading Libraries

```{r load-packages, include=FALSE}
library(readr)
library(here)
library(tidyverse)
library(ggthemes)
```


```{r}
moma <- read_csv("artworks-cleaned.csv")
```
# Know Your Data

## What is the first painting that was acquired?
```{r}
moma %>% 
  select(artist, title, year_acquired) %>% 
  arrange(year_acquired)
```
## What is the oldest painting?
```{r}
moma %>% 
  select(artist, title, year_created) %>% 
  arrange(year_created)
```

## How many artists?
```{r}
moma %>% 
  distinct(artist)
```

## Which artist has the most paintings?
```{r}
moma %>% 
  count(artist, sort = TRUE)
```

## How many paintings, by gender?
```{r}
moma %>% 
  count(artist_gender)
```

## How many artists, by gender?
```{r}
moma %>% 
  count(artist_gender, artist) %>% 
  count(artist_gender)
```

## In which years were the most paintings in the collection _acquired_?
```{r}
moma %>% 
  count(year_acquired, sort = TRUE)
```

## In which years were the most paintings in the collection _created_?
```{r}
moma %>% 
  count(year_created, sort = TRUE)
```

## What about the first painting by a solo female artist?
```{r}
moma %>% 
  filter(num_artists == 1 & n_female_artists == 1) %>% 
  select(title, artist, year_acquired, year_created) %>% 
  arrange(year_acquired)
```
# Basic Plotting!

## Year painted vs. year acquired
```{r}
ggplot(moma, aes(year_created, year_acquired)) +
  geom_point(alpha = .1, na.rm = TRUE) +
  geom_abline(intercept = c(0,0), colour = "red") +
  labs(x = "Year Painted", y = "Year Acquired") +
  ggtitle("MoMA Keeps Its Collection Current") 
```

# Exploring Painting Dimensions

## Challenge #4
```{r}
moma_dim <- moma %>% 
  filter(height_cm < 600, width_cm < 760) %>% 
  mutate(hw_ratio = height_cm / width_cm,
         hw_cat = case_when(
           hw_ratio > 1 ~ "taller than wide",
           hw_ratio < 1 ~ "wider than tall",
           hw_ratio == 1 ~ "perfect square"
         ))
library(ggthemes)
ggplot(moma_dim, aes(x = width_cm, y = height_cm, colour = hw_cat)) +
  geom_point(alpha = .5) +
  ggtitle("MoMA Paintings, Tall and Wide") +
  scale_colour_manual(name = "",
                      values = c("gray50", "#FF9900", "#B14CF0")) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) +
  labs(x = "Width", y = "Height") 

```

## Different colors
```{r}
library(ggthemes)
ggplot(moma_dim, aes(x = width_cm, y = height_cm, colour = hw_cat)) +
  geom_point(alpha = .5) +
  ggtitle("MoMA Paintings, Tall and Wide") +
  scale_colour_manual(name = "",
                      values = c("gray50", "#ee5863", "#6999cd")) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) +
  labs(x = "Width", y = "Height") 
```